---
title: "p8105_hw3_ly2565"
author: "Lin Yang"
date: "10/17/2021"
output: github_document
---

```{r, setup}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

## Load the instacart data.

```{r}
library(p8105.datasets)
data("instacart")
instacart
```

* The data set has `r nrow(instacart)` rows and `r ncol(instacart)` columns with variables: `r names(instacart)`. 

## Some questions related to this data set.
How many aisles are there, and which aisles are the most items ordered from?
```{r}
aisle_count = 
  instacart %>% 
  select(aisle) %>% 
  n_distinct()
aisle_count

aisle_most = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  mutate(aisle_rank = min_rank(desc(n_obs))) %>% 
  filter(aisle_rank < 3)
aisle_most
```

* There are `r aisle_count` aisles in total, and the aisles that the most items are ordered from are `r pull(aisle_most, aisle) `.

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.
```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  mutate(aisle = fct_reorder(factor(aisle), n_obs)) %>% 
  ggplot(aes(x = n_obs, y = aisle)) +
  geom_point() +
  labs(
    title = "Number of Items Ordered from Aisles with over 10000 Items",
    x = "Number of Items",
    y = "Aisles",
    caption = "Data from the Instacart Online Grocery Shopping Dataset 2017")
```














