---
title: "p8105_hw3_ly2565"
author: "Lin Yang"
date: "10/17/2021"
output: github_document
---

```{r, setup}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

## Load the instacart data.

```{r}
library(p8105.datasets)
data("instacart")
instacart
```

* The data set has `r nrow(instacart)` rows and `r ncol(instacart)` columns with variables: `r names(instacart)`. 

## Some questions related to this data set.
How many aisles are there, and which aisles are the most items ordered from:
```{r}
aisle_count = 
  instacart %>% 
  select(aisle) %>% 
  n_distinct()
aisle_count

aisle_most = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  mutate(aisle_rank = min_rank(desc(n_obs))) %>% 
  filter(aisle_rank < 3)
aisle_most
```

* There are `r aisle_count` aisles in total, and the aisles that the most items are ordered from are `r pull(aisle_most, aisle) `.

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered:
```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  mutate(aisle = fct_reorder(factor(aisle), n_obs)) %>% 
  ggplot(aes(x = n_obs, y = aisle)) +
  geom_point() +
  labs(
    title = "Number of Items Ordered from Aisles with over 10000 Items",
    x = "Number of Items",
    y = "Aisles",
    caption = "Data from the Instacart Online Grocery Shopping Dataset 2017")
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”:
```{r}
instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_items = n()) %>% 
  mutate(product_rank = min_rank(desc(n_items))) %>% 
  filter(product_rank < 4) %>% 
  group_by(aisle) %>% 
  arrange(-n_items) %>% 
  select(-product_rank) %>% 
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:
```{r}
mean_hour_dow =
  instacart %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )  
  
colnames(mean_hour_dow) = c("", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")
mean_hour_dow %>% 
  knitr::kable()
```











